\section{Informed Search}

\subsection{Heuristics}

\paragraph{Admissibility} An admissible heuristic never
overestimates the cost to reach the goal, that is it is
optimistic. A heuristic h is admissible if:
\begin{math}
  \forall\medspace node \medspace n, \medspace h(n) \leq h^*(n)
\end{math}
where $h^*$ is true cost to reach the goal from $n$.

In a map, the air distance between two points is an admissible
heuristic because it is by definition the shortest distance
between the two points.

\paragraph{Consistency}

\begin{math}
h(a) - h(b) \leq cost(a to b)
\end{math}

Which implies:
\begin{math}
h(a) \leq cost(a to b) + h(b)
\end{math}

\begin{table}[hb!]
\begin{tabular}{lL}
  Strategy & Implementation \\
  Greedy search &
  \multicolumn{1}{p{0.65\linewidth}}{
    Expands the node that appears to be closest to goal
  } \\
  $A^*$ &
  \multicolumn{1}{p{0.65\linewidth}}{
    Minimize the total estimated solution cost ($f$), that is
    the cost to reach node $n$ ($g$), and the cost to
    get from $n$ to the goal ($h$)
  }
\end{tabular}
\caption{Informed Search Strategies}
\label{tab:informed_search_strategies}
\end{table}

\subsection{A*}

\begin{itemize}
  \item Complete
  \item Time: exponential
  \item Space: keeps every node in memory
\end{itemize}


\subsection{A* optimality}

If the heuristic function used by A* is admissible, then A* is admissible. 

\paragraph{Sketch of proof} When A* terminates its search, 
it has found a path from start to goal whose actual cost is 
lower than the estimated cost of
any path from start to goal through any open node (the node's
$f$ value). 
When the heuristic is admissible, those estimates are
optimistic (not quite—see the next paragraph), so $A^{*}$ can
safely ignore those nodes because they cannot possibly lead
to a cheaper solution than the one it already has. 
In other words, $A^{*}$ will never overlook the
possibility of a lower-cost path from start to goal and so it
will continue to search until no such possibilities exist.

The actual proof is a bit more involved because the
$f$ values of open nodes are not guaranteed to
be optimistic even if the heuristic is admissible. This is
because the $g$ values of open nodes are not guaranteed to be
optimal, so the sum $g+h$ is not guaranteed to be optimistic.

