\section{Search Agents}

\paragraph{Applications}
\begin{itemize}
  \item Route-finding problem
  \item Travelling salesmane person
  \item VSLI layout: position million of components and
    connections on a chip to minimize area, shorten delays. Aim:
    put circuit components on a chip so as they don’t overlap and
    leave space to wiring which is a complex problem.
  \item Robot navigation
  \item Automatic assembly sequencing
  \item Protein folding
\end{itemize}

\subsection{Problem formulation}

\begin{itemize}
  \item Initial state: the state in which the agent starts
  \item States: All states reachable from the initial state by
    any sequence of actions (State space)
  \item Actions: possible actions available to the agent. At a
    state $s$, $Actions(s)$ returns the set of actions that can be
    executed in state $s$. (\textit{Action space})
  \item Transition model: A description of what each action does
    $Results(s, a)$
  \item Goal test: determines if a given state is a goal state
  \item Path cost: function that assigns a numeric cost to a path
    w.r.t. performance measure
\end{itemize}

\subsection{Search Space}

\begin{itemize}
  \item State space: a physical configuration
  \item Search space: an abstract configuration represented by a search tree or graph of possible solutions.
  \item Search tree: models the sequence of actions
    \begin{itemize}
      \item Root: initial state
      \item Branches: actions
      \item Nodes: results from actions. A node has: parent, children, depth, path cost, associated state in the state space.
    \end{itemize}
  \item Expand: A function that given a node, creates all children nodes
\end{itemize}

The search space is divided into three regions:
\begin{itemize}
  \item Explored (a.k.a. Closed List, Visited Set)
  \item Frontier (a.k.a. Open List, the Fringe)
  \item Unexplored.
\end{itemize}

\paragraph{Seach tree} 
\begin{itemize}
    \item A 'what if' tree of plans and their outcomes
    \item the start state is the root node
    \item Children correspond to successors
    \item Nodes show states, but correspond to plans that achieve those states
\end{itemize}
Each node in the search tree is an entire path in the state space graph


\subsection{Search Strategies}

\paragraph{Strategy evaluation} Strategies are evaluated along
the following dimensions:
\begin{itemize}
  \item Completeness: Does it always find a solution if one
    exists?
  \item Time complexity: Number of nodes generated/expanded
  \item Space complexity: Maximum number of nodes in memory
  \item Optimality: Does it always find a least-cost solution?
\end{itemize}

Time and space complexity: are measured in terms of:
\begin{itemize}
  \item $b$: maximum branching factor of the search tree (actions
    per state).
  \item $d$: depth of the solution
  \item $m$, or $D$: maximum depth of the state space (may be
    $\infty$).
\end{itemize}
